### 2021-09-18 plos_latex_template.tex
Rocío: major edits to rule 3 team including pay your team

Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 \ref{rule_inclusion} for more on that). 
But it will prevent from falling into the mistake of treating diversity as a checklist.
It will also guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference would translate into an even gender distribution in your speakers? 
Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way. 
%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).

@@

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
%i took the networking part to expand communication att.ast we can switch back afterwards

@@

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.

@@

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}) discussing the dimensions of diversity lacking in your community and event. To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for biases in its composition.

%liz: please check if this paragraph connects well with the whole rule. i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team. att: andrea

Gathering a diverse team will only work if there is real inclusion. People with disabilities often say: "Nothing about us without us"; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear. So, if you plan to organize something related to any community, region, or specific dimension of diversity, involve people from these groups.

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes far outweigh any inconveniences. %sara: inconveniences here doesn't sound good. I'd use: the positive outcomes are worth it
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. In addition, a diverse team plays an important role at creating a welcoming space because representation -seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes- is one of the best ways to create a sense of belonging (see Rule \ref{rule_inclusion}). This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”. 
%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general. Sara: maybe you can use this as we spoke relating team work here. And I think the working together part is not here but the following paragraph 
%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics

Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups. First, rethink power relationships inside the group and share power. Be committed to examine power relations inside your organizing team and challenge it during all the process.
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer coordinating positions, and let people from privileged groups step down. 
Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too.

And most importantly, take care of your team and prioritize their well-being! Check on them regularly and make sure everyone is comfortable. This represents a great amount of work and you'll need to support each other in the long run.

%Rocío: Yani's ref to have just in case:
% https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters

\section{Consciously unbias your spotlight roles}
\label{rule_spotlight}%sara: rule_unbias?

>>>>

Investing more in effort in the most excluded groups does not mean neglecting the others (see \textbf{Rule \ref{rule_inclusion}} for more on that). 
But it will prevent from falling into the mistake of treating diversity as a checklist.
It will also guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference would translate into an even gender distribution in your speakers? 
Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way. 
%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).

@@

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
%i took the networking part to expand communication att.ast we can switch back afterwards

@@

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.

@@

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition. 
%Rocío: I changed "biases" for "gaps". I'm sure other statisticians would've found the phrasing weird. To really understand what is meant by biases, the reader would have to go to rule unbias, and rules should be self-contained.
% Rocío: Idea of this paragraph: build a diverse team

Gathering a diverse team will only work if there is real inclusion. People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear. 
%liz, please check if this paragraph connects well with the whole rule. i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team. att: andrea
% Rocío: Idea of this paragraph: real inclusion when having diverse members

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}). 
% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least for now

%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”. 
%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general. Sara: maybe you can use this as we spoke relating team work here. And I think the working together part is not here but the following paragraph 
%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics

The following advice should apply to any kind of team, but are specially relevant when working with people from marginalized groups. First, rethink power relationships inside the group and share power. Be committed to examine power relations inside your organizing team and challenge it during all the process. % Rocío: Could we explain power relationship / share power ideas in more simple terms?
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person has their voice and enforce voices from systematically excluded groups.
Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. 
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.

Most importantly, take care of your team. 
Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
However, having a strong community as the only reward may be enough for those who are in more privileged position. 
Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. 
In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}). 
% Rocío: Idea of this paragraph: take care of your team.

\section{Consciously unbias your spotlight roles}
\label{rule_unbias}
